### 2020-06-02 writeup/intro.tex
sec 3
add overview
fred comments

%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning represents a powerful paradigm to solve complex tasks in computer vision \cite{chexnet17,ZSSGM18}, natural language processing \cite{GLUE,superglue} and many other areas \cite{ZY17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
By combining multiple information sources, it is possible to share all the information in the same model.
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
The performance of multi-task learning depends on the relationship of the information sources \cite{C97}.
When the information sources are heterogeneous, negative transfer--where multi-task learning (MTL)  performs worse than single-task learning (STL)--has often been observed \cite{PY09,AP16,BS17}.
While many empirical approaches are proposed to mitigate negative transfer \cite{ZY17}, a precise understanding of when negative transfer occurs remains elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.
In this work, we show that negative transfer is determined by the bias-variance tradeoff of multi-task learning estimators for the high-dimensional linear regression setting. % for learning from multiple linear regression tasks.
We develop new tools to analyze the tradeoff and explain negative transfer precisely.
%Inspired by the theory, we propose an incremental training schedule to improve multi-task training.
%We consider a setting where the target task has limited labeled data and show
%On the other hand, unless the structures across task data are well-understood, applying multi-task learning on several different datasets often result in suboptimal models (or negative transfer in more technical terms).

@@

\begin{theorem}[Informal]\label{lem_cov_shift_informal}
	In the setting of two tasks, let $a_1, a_2$ be fixed values solved from deterministic equations of $\rho_1, \rho_2$ and the spectrum of $\Sigma_1^{1/2}\Sigma_2^{-1}\Sigma_1^{1/2}$.
	%let $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in descending order.
	For any constant $\e>0$, w.h.p. over the randomness of $X_1, X_2$, we have that
	\begin{align*}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{\rho_1+\rho_2}\cdot \frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2} +\bigo{{p^{-1/2+\epsilon}}}.
	\end{align*}
\end{theorem}

%Concretely, we show a tight bound on the trace of $(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}$, which
Theorem \ref{lem_cov_shift_informal} extends a well-known result for the single-task setting when $X_1, \rho_1, a_1 = 0$ \cite{S07}.
The concentration error of our result, which is on the order of $O(p^{-1/2+\varepsilon})$, is nearly optimal.
%We shall assume that each task data follows a linear model, i.e. $y_i = X_i \beta_i + \varepsilon_i$, $1\le i\le k$.
%Here $\beta_i\in\real^p$ is the model parameter for the $i$-th task.
%Each row of $X_i\in\real^{n_i\times p}$ is assumed to be drawn i.i.d. from a fixed
%distribution with covariance matrix $\Sigma_i$.
%We say there is negative transfer if the prediction loss of $\hat{\beta}_t^{\MTL}$  is larger than that of $\hat{\beta}_t^{\STL}$, or positive transfer otherwise.
Based on Theorem \ref{lem_cov_shift_informal}, we provide an analysis of the bias-variance tradeoff of $\hat{\beta}_t^{\MTL}$ for two settings:
(i) two tasks with general covariance matrice; (ii) many tasks with the same features (e.g. images).
We extend our result to the transfer learning
%in the setting of high-dimensional linear regression.
%by pooling source task representations into the shared body of the hard parameter sharing architecture, following
setting of Taskonomy by Zamir et al. \cite{ZSSGM18}.
We prove that the bias of the transfer learning estimator is given by the projection of $\beta_t$ to the orthogonal subspace spanned by $\set{\beta_i}_{i=1}^{t-1}$.
These results are described more precisely in Section \ref{sec_main}.

>>>>

%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning represents a powerful paradigm to solve complex tasks in computer vision \cite{chexnet17,ZSSGM18}, natural language processing \cite{GLUE,superglue} and many other areas \cite{ZY17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
%By combining multiple information sources, it is possible to share all the information in the same model.
In many settings, multiple information sources are available to help solve a particular task.
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
The performance of multi-task learning depends on the relationship between the information sources and the task \cite{C97}.
When the information sources are heterogeneous, negative transfer--where multi-task learning (MTL)  performs worse than single-task learning (STL)--has often been observed \cite{PY09,AP16,BS17}.
While many empirical approaches are proposed to mitigate negative transfer \cite{ZY17}, a precise understanding of when negative transfer occurs remains elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.
In this work, we show that negative transfer is determined by the bias-variance tradeoff of multi-task learning estimators for the high-dimensional linear regression setting. % for learning from multiple linear regression tasks.
We develop new tools to analyze the tradeoff and explain negative transfer precisely.
%Inspired by the theory, we propose an incremental training schedule to improve multi-task training.
%We consider a setting where the target task has limited labeled data and show
%On the other hand, unless the structures across task data are well-understood, applying multi-task learning on several different datasets often result in suboptimal models (or negative transfer in more technical terms).

@@

\begin{theorem}[Informal]\label{lem_cov_shift_informal}
	In the setting of two tasks, let $a_1, a_2$ be fixed values solved from deterministic equations of $\rho_1, \rho_2$ and the spectrum of $\Sigma_1^{1/2}\Sigma_2^{-1}\Sigma_1^{1/2}$.
	%let $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in descending order.
	For any constant $\e>0$, w.h.p. over the randomness of $X_1, X_2$, we have that
	\begin{align}\label{eq_introX1X2}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{\rho_1+\rho_2}\cdot \frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2} +\bigo{{p^{-1/2+\epsilon}}}.
	\end{align}
\end{theorem}

%Concretely, we show a tight bound on the trace of $(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}$, which
Theorem \ref{lem_cov_shift_informal} extends a well-known result for the single-task setting when $X_1, \rho_1, a_1 = 0$ \cite{S07}.
The concentration error of our result, which is on the order of $O(p^{-1/2+\varepsilon})$, is nearly optimal.
%We shall assume that each task data follows a linear model, i.e. $y_i = X_i \beta_i + \varepsilon_i$, $1\le i\le k$.
%Here $\beta_i\in\real^p$ is the model parameter for the $i$-th task.
%Each row of $X_i\in\real^{n_i\times p}$ is assumed to be drawn i.i.d. from a fixed
%distribution with covariance matrix $\Sigma_i$.
Based on Theorem \ref{lem_cov_shift_informal}, we provide an analysis of the bias-variance tradeoff of $\hat{\beta}_t^{\MTL}$ for two settings:
(i) two tasks with general covariance matrice; (ii) many tasks with the same features (e.g. images).
We extend our result to the transfer learning
%in the setting of high-dimensional linear regression.
%by pooling source task representations into the shared body of the hard parameter sharing architecture, following
setting of Taskonomy by Zamir et al. \cite{ZSSGM18}.
We prove that the bias of the transfer learning estimator is given by the projection of $\beta_t$ to the orthogonal subspace spanned by $\set{\beta_i}_{i=1}^{t-1}$.
These results are described more precisely in Section \ref{sec_main}.

### 2020-06-02 writeup/setup.tex
fred comments

\textbf{Notations.}
When there is no ambiguity, we drop the subscript $t$ from $\te_t(\hat{\beta}_t^{\MTL})$ to $\te(\hat{\beta}_t^{\MTL})$ for simplicity.
We refer to the first task as the source task when there are only two tasks.
%We call $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ the covariate shift matrix.
We say there is negative transfer if the prediction loss of $\hat{\beta}_t^{\MTL}$ is larger than that of $\hat{\beta}_t^{\STL}$, or positive transfer otherwise.
For a matrix $X\in\real^{p_1\times p_2}$, let $\lambda_{\min}(X)$ denote its minimum singular value.

@@

Equation \eqref{eq_te_model_shift} corresponds to the bias of $\hat{\beta}_t^{\MTL}$, which captures how similar $\beta_1$ and $\beta_2$ are.
Hence, the bias of $\hat{\beta}_t^{\MTL}$ introduces a negative effect from adding the source labels.
Equation \eqref{eq_te_var} corresponds to the variance of $\hat{\beta}_t^{\MTL}$, which is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
%This part introduces a positive variance reduction effect from adding the source labels.
Hence, whether $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ is determined precisely by the bias-variance tradeoff!
%(i) the negative effect from model shift bias.
%(ii) the positive effect from variance reduction;

@@

\textbf{Multi-task learning.}
We provide a sharp analysis of the bias-variance tradeoff for two settings:
(i) two tasks with general covariance matrices;
(ii) any number of tasks that have the same features, i.e. $X_i = X_j$ for any $i\neq j$.
The latter setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
We state our result for two tasks as follows.
%The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
%We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
%Based on the result, we show how to determine positive versus negative transfer as follows.

\begin{theorem}[Informal statement of Theorem \ref{thm_model_shift}]\label{thm_main_informal}
	For the setting of two tasks, let $M=\Sigma_1^{1/2}\Sigma_2^{-1/2}$, $\delta > 0$ be a desired error margin and $\rho_1 \gtrsim \frac{1}{\delta^2}\cdot \lambda_{\min}(M)^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$, where $\lambda_{\min}(M)$ is the smallest singular value of $M_1$.
	There exists two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on $\set{\hat{v}, \Sigma_1, \Sigma_2, \rho_1, \rho_2, \beta_1, \beta_2}$ such that
	\squishlist
		\item If $\Delta_{\bias} - \Delta_{\vari} < -\delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\bias} - \Delta_{\vari} > \delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}

Theorem \ref{thm_main_informal} shows that as the source task data size increases, we obtain a sharp transition from positive transfer to negative transfer determined by $\Delta_{\bias} - \Delta_{\vari}$.
%determined by the covariate shift matrix and the model shift.
%The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of the threshold can be complex (as is preivous generalization bounds for MTL), they admit interpretable forms for simplified settings.
This will be the focus of Section \ref{sec_insight}.

@@

%The proof of Theorem \ref{thm_main_informal} is presented in Appendix \ref{app_proof_main}.
We have a similar theorem for more than two tasks but with the matrix $X_i=X$, and it is deferred to Appendix \ref{app_proof_many_tasks}.

>>>>

\textbf{Notations.}
When there is no ambiguity, we drop the subscript $t$ from $\te_t(\hat{\beta}_t^{\MTL})$ to $\te(\hat{\beta}_t^{\MTL})$ for simplicity.
We refer to the first task as the source task when there are only two tasks.
%We call $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ the covariate shift matrix.
We say there is negative transfer if the prediction loss of $\hat{\beta}_t^{\MTL}$ is larger than that of $\hat{\beta}_t^{\STL}$, or positive transfer otherwise.
For a matrix $X\in\real^{p_1\times p_2}$, let $\lambda_{\min}(X)$ denote its minimum singular value.
Let $\norm{X}$ denote the spectral norm of $X$.

@@

Equation \eqref{eq_te_model_shift} corresponds to the bias of $\hat{\beta}_t^{\MTL}$, which captures how similar $\beta_1$ and $\beta_2$ are.
Hence, the bias of $\hat{\beta}_t^{\MTL}$ introduces a negative effect from adding the source labels.
Equation \eqref{eq_te_var} corresponds to the variance of $\hat{\beta}_t^{\MTL}$, which is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
%This part introduces a positive variance reduction effect from adding the source labels.
Hence, whether $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ is determined precisely by the tradeoff between the negative effect of the bias term and the positive effect of the variance term!
%(i) the negative effect from model shift bias.
%(ii) the positive effect from variance reduction;

@@

\textbf{Multi-task learning.}
We provide a sharp analysis of the bias-variance tradeoff for two settings:
(i) two tasks with general covariance matrices;
(ii) any number of tasks that have the same features, i.e. $X_i = X_j$ for any $i\neq j$.
The latter setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
We state our result for two tasks as follows.
%The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
%We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
%Based on the result, we show how to determine positive versus negative transfer as follows.
%, where $\lambda_{\min}(M)$ is the smallest singular value of $M_1$
\begin{theorem}[Informal statement of Theorem \ref{thm_model_shift}]\label{thm_main_informal}
	For the setting of two tasks, let $M=\Sigma_1^{1/2}\Sigma_2^{-1/2}$, $\delta > 0$ be a desired error margin and $\rho_1 \gtrsim \frac{1}{\delta^2}\cdot \lambda_{\min}(M)^{-4} \norm{\Sigma_1} \max(\norm{\beta_1}^2, \norm{\beta_2}^2)$.
	Let $\rho_2 > 1$ be a fixed value.
 	There exists two deterministic functions $\Delta_{\bias}$ and $\Delta_{\vari}$ that only depend on $\set{\hat{v}, \Sigma_1, \Sigma_2, \rho_1, \rho_2, \beta_1, \beta_2}$ such that
	\squishlist
		\item If $\Delta_{\bias} - \Delta_{\vari} < -\delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\bias} - \Delta_{\vari} > \delta$, then w.h.p. over the randomness of $X_1, X_2$, we have $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}

Theorem \ref{thm_main_informal} applies to settings where large amounts of source task data is available but the target data size is small.
For such settings, we obtain a sharp transition from positive transfer to negative transfer determined by $\Delta_{\bias} - \Delta_{\vari}$.
%determined by the covariate shift matrix and the model shift.
%The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of the threshold can be complex (as is preivous generalization bounds for MTL), they admit interpretable forms for simplified settings.
This will be the focus of Section \ref{sec_insight}.

@@

%The proof of Theorem \ref{thm_main_informal} is presented in Appendix \ref{app_proof_main}.
A similar result for the second setting can be found in Appendix \ref{app_proof_many_tasks}.

### 2020-06-02 writeup/transfer.tex
fred comments

We provide precise explanations to the phenomena of negative transfer in multi-task learning.
We explain from three perspectives, including \textit{task similarity}, \textit{data size} and \textit{covariate shift}.
We show how negative transfer occurs as a result in a simplified model and describe the algorithmic consequence.

@@

\textbf{Algorithmic consequence.}
As an extension of Proposition \ref{prop_dist_transition}, we observe that MTL is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
If $\sigma_1$ is too large, the source task provides a negative transfer to the target.
If $\sigma_1$ is small, the source task is more helpful.
We leave the result to the Appendix in Proposition \ref{prop_var_transition}.
Inspired by the observation, we propose a single-task based metric to help understand MTL results using STL results.
\squishlist
	\item For each task, we train a single-task model.
	Let $z_s$ and $z_t$ be the prediction accuracy of each task, respectively.
	Let $\tau\in(0, 1)$ be a fixed threshold.
	\item If $z_s - z_t > \tau$, then we predict that there will be positive transfer when combining the two tasks using MTL.
	If $z_s - z_t < -\tau$, then we predict negative transfer.
\squishend

@@

\begin{proposition}[Source/task data ratio]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 110$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} < \frac{1}{\gamma_+} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1) - \gamma_+^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} > \frac{1}{\gamma_-} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1.5) - \gamma_{-}^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} describes the bias-variance tradeoff in terms of the data ratio $n_1 / n_2$.
From Figure \ref{fig_size}, we see that our result is able to explain the transition from positive to negative transfer.
There are several unexplained observations near $y = 0$ caused by the errors $\gamma_-, \gamma_+$.
The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_32}.

\textbf{Connection to Taskonomy.} We use our tools to explain a key result of Taskonomy by Zamir et al.'18 \cite{ZSSGM18}, which shows that MTL can reduce the amount of labeled data needed to achieve comparable performance to STL.
Formally, let the data efficiency ratio be the largest factor $x$ such that the total number of labeled datapoints needed for solving all tasks can be reduced by $x$ while keeping the same performance compared to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using only $x \cdot n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
Using our tools, we show that in the isotropic model, the data efficiency ratio is
equal to \[ \frac{1}{\rho_1 + \rho_2} {+ \frac{1}{(\rho_1 +\rho_2)(\rho_1^{-1} + \rho_2^{-1} - \Theta(\Psi(\beta_1, \beta_2)))}}. \]
In addition to $\rho_1,\rho_2$, the data efficiency ratio also scales with $\Psi(\beta_1, \beta_2)$, the model distance between the tasks.
The result can be found in Appendix \ref{app_proof_32}.

\textbf{Algorithmic consequence.} An interesting consequence of Proposition \ref{prop_data_size} is that $L(\hat{\beta}_t^{\MTL})$ is not monotone in $\rho_1$.
In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
Based on this insight, we propose an incremental optimization schedule to improve MTL training efficiency.
Algorithm \ref{alg_inc_train} describes the procedure.

@@

By applying Lemma \ref{lem_cov_shift_informal}, we find that when $n_1 / n_2$ is large, having no covariate shift is the optimal choice provided that the determinant of $M^{\top}M$ is bounded.
To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}[Covariate shift]\label{prop_covariate}
	In the setting described above, assume that $\rho_1, \rho_2>1$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

>>>>

We provide precise explanations to the phenomena of negative transfer in multi-task learning.
We explain from three perspectives, including \textit{task similarity}, \textit{data size} and \textit{covariate shift}.
We show how negative transfer occurs by varying task similarity or data size.
Then we show that when source task data size becomes large, covariate shift causes more negative effects.

@@

\textbf{Algorithmic consequence.}
We can in fact extend the result to the cases where the noise variance is different.
In this case, we will see that MTL is particularly effective.
%As an extension of Proposition \ref{prop_dist_transition}, we observe that MTL is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Concretely, suppose the noise variance $\sigma_1^2$ of task $1$ differs from the noise variance $\sigma_2^2$ of task $2$.
If $\sigma_1^2$ is too large, the source task provides a negative transfer to the target.
If $\sigma_1^2$ is small, the source task is more helpful.
We leave the result to the Appendix in Proposition \ref{prop_var_transition}.
Inspired by the observation, we propose a single-task based metric to help understand MTL results using STL results.
\squishlist
	\item For each task, we train a single-task model.
	Let $z_s$ and $z_t$ be the prediction accuracy of each task, respectively.
	Let $\tau\in(0, 1)$ be a fixed threshold.
	\item If $z_s - z_t > \tau$, then we predict that there will be positive transfer when combining the two tasks using MTL.
	If $z_s - z_t < -\tau$, then we predict negative transfer.
\squishend

@@

\begin{proposition}[Source/target data size]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 110$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} < \frac{1}{\gamma_+} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1) - \gamma_+^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} > \frac{1}{\gamma_-} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1.5) - \gamma_{-}^{-1}}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} describes the bias-variance tradeoff in terms of the data ratio $n_1 / n_2$.
From Figure \ref{fig_size}, we see that our result is able to explain the transition from positive to negative transfer.
There are several unexplained observations near $y = 0$ caused by the errors $\gamma_-, \gamma_+$.
The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_32}.

\textbf{Connection to Taskonomy.} We use our tools to explain a key result of Taskonomy by Zamir et al.'18 \cite{ZSSGM18}, which shows that MTL can reduce the amount of labeled data needed to achieve comparable performance to STL.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(x)$ denote the estimator trained using $R \cdot n_i$ datapoints from every task.
The data efficiency ratio roughly scales as
\[ \argmin_{x\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
For example, the data efficiency ratio is $1$ if there is negative transfer.
Using our tools, we show that in the isotropic model, the data efficiency ratio is
equal to \[ \frac{1}{\rho_1 + \rho_2} {+ \frac{1}{(\rho_1 +\rho_2)(\rho_1^{-1} + \rho_2^{-1} - \Theta(\Psi(\beta_1, \beta_2)))}}. \]
Compared with Proposition \ref{prop_dist_transition}, we see that when $\Psi(\beta_1, \beta_2)$ is smaller than $\rho_1^{-1} + \rho_2^{-1}$, the transfer is positive.
Hence the data efficiency ratio quantifies how effective is the positive transfer using MTL.
%In addition to $\rho_1,\rho_2$, the data efficiency ratio also scales with $\Psi(\beta_1, \beta_2)$, the model distance between the tasks.
The precise statement can be found in Appendix \ref{app_proof_32}.

\textbf{Algorithmic consequence.} An interesting consequence of Proposition \ref{prop_data_size} is that $L(\hat{\beta}_t^{\MTL})$ is not monotone in $\rho_1$.
In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
Based on this insight, we propose an incremental optimization schedule to improve MTL training efficiency.
\squishlist
	\item We divide the source task data into $S$ batches.
	For $S$ rounds, we incrementally add the source task data by adding one batch at a time.
	\item After training $T$ epochs, if the validation accuracy becomes worse than the previous round's result, we terminate.
	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\squishend

@@

%By applying Lemma \ref{lem_cov_shift_informal}, we find that when $n_1 / n_2$ is large, having no covariate shift is the optimal choice provided that the determinant of $M^{\top}M$ is bounded.
To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, let $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M$ in decreasing order.
Let $\mu_{\min} < \mu < \mu_{\max}$ are fixed values that do not grow with $p$.
Consider the following bounded set
\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \prod_{i=1}^p \lambda_i \le \mu^p, \mu_{\min} \le \lambda_p, \lambda_1 \le \mu_{\max}},
\end{align*}
%	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.

\begin{proposition}[Covariate shift]\label{prop_covariate}
	Assume that $\Psi(\beta_1, \beta_2) = 0$ and $\rho_1, \rho_2>1$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
The following proposition shows that when $n_1 \gg n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized when there is no covariate shift between source and target task.
The proof of Proposition \ref{prop_covariate} is left to Appendix \ref{app_proof_33}.
%Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
